---
title: "Chapter 2"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggpubr)
```

# TESTS AND CONFIDENCE INTERVALS FOR MEANS

## 2.1 “Test for Mean μ When Standard Deviation ϭ ls Known”

### 2.1.2 Normal Test

R을 이용해서 테스트를 해봄.

```{r}
n <- 26 # 테스트한 로켓 숫자
m = 2000  # 평균거리
s = 120   # 거리 편차
alpha = .05  # 유의수준
ntest <- 100000  # 가상 실험 횟수, 많을 수록 정확
am = 1940 # 대립 평균 (샘플로 부터 얻은 평균)
```

가상 실험을 수행했을때,test를 계속 수행하면,
결국, 

```{r}
res <- list()
for (i in 1:ntest) {
  # 대립평균, 동일 편차로 sampling
  new_sample <- rnorm(n, mean = am, sd = s)
  # one-sided t-test에서 p-value
  test_res <- t.test(new_sample, mu = m, alternative ="less")
  res[[i]] <- test_res$p.value
}
res <- unlist(res)
```

```{r}
df <- tibble(distance = res)
df
```

```{r}
df %>% 
  ggplot(aes(x = distance)) +
  geom_density() +
  geom_vline(xintercept = 0.05, color = "red")
```

```{r}
(sum(res <= 0.05) / ntest )
```

```{r}
?t.test
```



### 2.1.3 Confidence Intervals

```{r}
m = 1.215
s = 0.01
alpha = 0.05
n = 4

ntest = 100000

rs <- list()
for (i in 1:ntest) {
  r <- rnorm(n, mean = m, sd = s)
  rs[[i]] <- mean(r)
}
```


```{r}
hist(unlist(rs))
```


```{r}
df <- tibble(rs = unlist(rs))
```

```{r}
df %>%
  ggplot(aes(x = rs)) +
  geom_density() + 
  #geom_vline(xintercept = 0.05, color = "red")
```

```{r}
df %>%
  filter(rs > 1.205, rs < 1.225) %>%
  count()
```



## 2.2 Test for mu when sigma is unknown

### 2.2.2 The t Test

```{r}

xi <- c(2.021, 2.002, 2.001,
        2.005, 1.990, 1.990,
        2.009, 1.983, 1.987)
alpha <- 0.05
mu <- 2.000
yi <- 1000 * (xi - 2)
yi
```


```{r}
t.test(xi, mu = 2)
```



```{r}
plot(extra ~ group, data = sleep
)
```


```{r}
sleep
```

```{r}
sleep %>% pivot_wider(names_from = group, values_from = extra)
```


```{r}
t.test(extra ~ group, data = sleep)
```











### 2.2.3
```{r}
n = 10
x = 42
s = 2.1
```


```{r}
res <- list()
for (i in 1:100000) {
  res[[i]] <- mean(rnorm(10, mean = 42, sd = 2.1))
}
df <- tibble(res = unlist(res))
```

```{r}
a <- rnorm(10, mean = 42, sd = 2.1)

```


```{r}
df
```

```{r}
df %>% 
  ggplot(aes(x = res)) +
  geom_density()
```

```{r}
quantile(df$res, probs = c(0.025, 0.975))
```

## 2.3

### 2.3.2 Test

```{r}
# must be p = 0.007
binom.test(3, 40, p = 0.01)

# must be p = 0.061
binom.test(2, 40, p = 0.01)

```


## 2.4 Test for μ - μ2 When σ1, and σ2 Are Known”


### 2.4.2 Normal Test


```{r}
x1 <- c(63.12, 63.57, 62.81, 64.32, 63.76)
x2 <- c(62.54, 63.21, 62.38, 62.06)
x_sd <- 0.6

mean(x1)
mean(x2)
```

```{r}
z <-  (mean(x1) - mean(x2)) / sqrt(x_sd^2 / length(x1) + x_sd^2 / length(x2))
z
```



### 2.5.2 Test for Paired Observations

```{r}
##install.packages("BSDA")
```


```{r}
A <- c(4.4, -1.4, 3.2, 0.2, -5, 0.3, 1.2, 2.2, 1.3, -0.7)
B <- c(3.2, 7.7, 6.4, 2.7, 3.1, 0.6, 2.6, 2.2, 2.2, 0.9)
mean(A)
mean(B)
```

```{r}
library(BSDA)
SIGN.test(A, B)
```

```{r}
library(tidyverse)
# pooled standard deviation from wikipedia

d1 <- tibble(x = 1, y = c(31, 30, 29))
d2 <- tibble(x = 2, y = c(42, 41, 40, 39))
d3 <- tibble(x = 3, y = c(31, 28))
d4 <- tibble(x = 4, y = c(23, 22, 21, 19, 18))
d5 <- tibble(x = 5, y = c(21, 20, 19, 18, 17))

df <- bind_rows(d1, d2, d3, d4, d5)
df
```

```{r}
df2 <- df %>%
  group_by(x) %>%
  summarize(
    n = n(),
    mean = mean(y),
    var = var(y),
    sd = sd(y)
  )
df2
```

```{r}
df2 %>% 
  mutate(
    n_ = n - 1,
    v = n_ * var
  ) %>%
  summarise(sum = sum(v), nsum = sum(n_), sp_sq = sqrt((sum / nsum)))
```

```{r}
mean(df2$sd)
```

### 2.5.3 

```{r}
df3 <- tribble(
  ~m1, ~m2,
  98.06, 94.73,
  98.07, 95.02,
  98.93, 94.36,
  98.03, 96.16,
  98.50, 94.50,
  98.60, 94.82,
  99.67, 93.75,
  98.51, 93.90,
  98.06, 95.31,
  99.01, 95.21
)
df3
```

```{r}
df3 %>%
  summarize_all(mean)
```


```{r}
## example Two Samle t-test
## note that var.equal = TRUE
t.test(df3$m1, df3$m2, mu = 3, paired = FALSE, var.equal = TRUE)
```



### 2.5.4 Test of μ1 - μ2 = 0 for Two Independent Samples When σ1≠σ2

```{r}
x2 <- c(
  1472, 1315, 1984, 1900, 1646,
  1709, 1780, 1571, 1681,1454,
  1799, 1372, 1420, 1612, 1546,
  1255, 1818, 1357, 1412, 1775,
  1850, 1417, 2057, 1468, 1694,
  1776, 1456, 1489, 1618,1544
)

x1 <- c(
  1747, 1618, 1521, 1137, 1374,
  1325, 1821, 2351, 1883, 1613,
  1843, 1796, 1507, 1387, 725, 
  1041, 1652, 1595, 1679, 1557, 
  1206, 192, 1025, 813, 971
)
```


```{r}
## example Two Samle t-test
## note that var.equal = FALSE
t.test(x2, x1, mu = 0, paired = FALSE, var.equal = FALSE)
```


### 2.8 Detection of a trend in a set of means, standard deviations unknown but assumed equal

```{r}
week <- c(4, 8, 12, 16, 20, 24, 27.5, 30, 35)
glucose <- c(43.14, 42.16, 45, 44, 47.86, 48.15, 50.10, 49.05, 50.15)
df <- tibble(
  week = c(4, 8, 12, 16, 20, 24, 27.5, 30, 35),
  glucose = c(43.14, 42.16, 45, 44, 47.86, 48.15, 50.10, 49.05, 50.15)
)
df
```

```{r}
df <- df %>%
  mutate(diff = lead(glucose) - glucose)
df

```


```{r}
var(df$diff, na.rm = TRUE) / var(glucose)
```





```{r}
a <- sum(df$diff^2, na.rm = TRUE) #/ (length(df$diff))
b <- sum(df$glucose^2) - sum(df$glucose)^2/length(df$glucose)
a
b
a / b
```


```{r}
a2 <- sum(df$diff^2, na.rm = TRUE) / (length(df$diff) - 1)
b2 <- sum((df$glucose - mean(df$glucose))^2) / (length(df$glucose) - 1)
a2
b2
a2 / b2
```

```{r}
# na.rm은 사전에 제거하므로 n 
a3 <- var(df$diff, na.rm = TRUE)
b3 <- var(df$glucose)
a3
b3
a3 / b3
```

```{r}
var(df$diff, na.rm = TRUE)
```


```{r}
var(df$diff[1:8])
```





























